---
title: "COVID-19 in the CA Bay Area"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(whisker)
library(RCurl)

# read in COVID-19 data
u <- getURL("https://raw.githubusercontent.com/datadesk/california-coronavirus-data/master/latimes-county-totals.csv")
covid <- read.csv(text = u)

# filter on Bay Area counties
bay <- filter(covid, 
              county %in% c("Alameda","Contra Costa","Marin","Monterey","Napa","San Francisco",
                           "San Mateo","Santa Clara","Santa Cruz","Solano","Sonoma") & confirmed_cases > 0
)

# clean up issue caused by confirmed_cases = 0 in June
bay$new_confirmed_cases[bay$date == "2020-06-04" & bay$county == "Alameda"] <- NA
bay$new_confirmed_cases[bay$date == "2020-06-05" & bay$county != "Alameda" & bay$county != "Contra Costa"] <-  NA

# format date
bay$date <- as.Date(bay$date, format = "%Y-%m-%d")

# read in population data
pop <- read_csv("pop_bayarea.csv")

# add county populations to COVID-19 data
# normalize infected total by 100K residents
bay <- inner_join(bay, pop) %>%
  mutate(confirmed_cases_norm = confirmed_cases*100000/population)

# separate data for each county
ac <- filter(bay, county == "Alameda")
cc <- filter(bay, county == "Contra Costa")
mc <- filter(bay, county == "Marin")
mo <- filter(bay, county == "Monterey")
nc <- filter(bay, county == "Napa")
sf <- filter(bay, county == "San Francisco")
sm <- filter(bay, county == "San Mateo")
scc <- filter(bay, county == "Santa Clara")
scz <- filter(bay, county == "Santa Cruz")
sol <- filter(bay, county == "Solano")
son <- filter(bay, county == "Sonoma")

# Whisker template for total cases (reusing for diff counties)
template_tot <-
'
{{county_abbv}}_tot <- ggplot({{county_abbv}}, aes(x = date,
                                 y = confirmed_cases,
                                 text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                               "<b>Total cases: </b>", confirmed_cases, "<br>",
                                               "<b>Additional cases today: </b>", new_confirmed_cases),
                                 group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  theme_minimal(base_size = 14) +
  geom_point(color = "{{county_color}}") +
  geom_line(color = "{{county_color}}")

ggplotly({{county_abbv}}_tot, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
'

# Whisker template for new cases (reusing for diff counties)
template_new <-
'
{{county_abbv}}_new <- ggplot(filter({{county_abbv}}, !is.na(new_confirmed_cases)), aes(x = date,
                                 y = new_confirmed_cases,
                                 text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                               "<b>Additional cases today: </b>", new_confirmed_cases),
                                 group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  theme_minimal(base_size = 14) +
  geom_col(fill = "{{county_color}}")

ggplotly({{county_abbv}}_new, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
'

# Whisker template for deaths (reusing for diff counties)
template_deaths <-
'
{{county_abbv}}_deaths <- ggplot({{county_abbv}}, aes(x = date,
                                    y = deaths,
                                    text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                                  "<b>Total deaths: </b>", deaths, "<br>",
                                                  "<b>Additional deaths today: </b>", new_deaths),
                                    group = 1)) +
  xlab("") +
  ylab("Deaths") +
  theme_minimal(base_size = 14) +
  geom_point(data = filter({{county_abbv}}, new_deaths > 0), color = "gray50") + 
  geom_line(color = "gray50")

ggplotly({{county_abbv}}_deaths, tooltip = "text") %>%
  config(displayModeBar = FALSE) %>%
  layout(hoverlabel = list(align = "left"))
'

# metadata for each county
ac_data <- list(county_abbv = "ac", county_color = "#a6cee3")
cc_data <- list(county_abbv = "cc", county_color = "#1f78b4")
mc_data <- list(county_abbv = "mc", county_color = "#b2df8a")
mo_data <- list(county_abbv = "mo", county_color = "#33a02c")
nc_data <- list(county_abbv = "nc", county_color = "#fb9a99")
sf_data <- list(county_abbv = "sf", county_color = "#e31a1c")
sm_data <- list(county_abbv = "sm", county_color = "#fdbf6f")
scc_data <- list(county_abbv = "scc", county_color = "#ff7f00")
scz_data <- list(county_abbv = "scz", county_color = "#cab2d6")
sol_data <- list(county_abbv = "sol", county_color = "#6a3d9a")
son_data <- list(county_abbv = "son", county_color = "#ffff99")

```

All
=====================================
- Numbers from [L.A. Times coronavirus data](https://github.com/datadesk/california-coronavirus-data/blob/master/latimes-county-totals.csv)
- Populations for normalization based on 2019 population estimates provided by [U.S. Census Bureau QuickFacts](https://www.census.gov/quickfacts/fact/table/US/PST045219)

Column {.tabset}
-------------------------------------
### Original Data, Linear Scale
```{r}
bay_total_chart <- ggplot(bay, aes(x = date,
                                   y = confirmed_cases,
                                   color = county,
                                   text = paste0("<b>County: </b>", county, "<br>",
                                                 "<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                                 "<b>Total cases: </b>", confirmed_cases, "<br>",
                                                 "<b>Additional cases today: </b>", new_confirmed_cases),
                                   group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Paired",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "Marin",
                            "Monterey",
                            "Napa",
                            "San Francisco",
                            "San Mateo",
                            "Santa Clara",
                            "Santa Cruz",
                            "Solano",
                            "Sonoma"))

ggplotly(bay_total_chart, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

### Original Data, Log Scale

```{r}
bay_total_log <- ggplot(bay, aes(x = date,
                                 y = confirmed_cases,
                                 color = county,
                                 text = paste0("<b>County: </b>", county, "<br>",
                                               "<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                               "<b>Total cases: </b>", confirmed_cases, "<br>",
                                               "<b>Additional cases today: </b>", new_confirmed_cases),
                                 group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  scale_y_continuous(trans = "log10")  +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Paired",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "Marin",
                            "Monterey",
                            "Napa",
                            "San Francisco",
                            "San Mateo",
                            "Santa Clara",
                            "Santa Cruz",
                            "Solano",
                            "Sonoma"))

ggplotly(bay_total_log, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

### Normalized Data, Linear Scale

```{r}
bay_norm_total_chart <- ggplot(bay, aes(x = date,
                                   y = confirmed_cases_norm,
                                   color = county,
                                   text = paste0("<b>County: </b>", county, "<br>",
                                                 "<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                                 "<b>Total cases (per 100K residents): </b>",
                                                 format(round(confirmed_cases_norm, 2), nsmall = 2)),
                                   group = 1)) +
  xlab("") +
  ylab("Confirmed cases per 100,000 county residents") +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Paired",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "Marin",
                            "Monterey",
                            "Napa",
                            "San Francisco",
                            "San Mateo",
                            "Santa Clara",
                            "Santa Cruz",
                            "Solano",
                            "Sonoma"))

ggplotly(bay_norm_total_chart, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

### Normalized Data, Log Scale

```{r}
bay_norm_total_log <- ggplot(bay, aes(x = date,
                                   y = confirmed_cases_norm,
                                   color = county,
                                   text = paste0("<b>County: </b>", county, "<br>",
                                                 "<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                                 "<b>Total cases (per 100K residents): </b>",
                                                 format(round(confirmed_cases_norm, 2), nsmall = 2)),
                                   group = 1)) +
  xlab("") +
  ylab("Confirmed cases per 100,000 county residents") +
  scale_y_continuous(trans = "log10")  +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Paired",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "Marin",
                            "Monterey",
                            "Napa",
                            "San Francisco",
                            "San Mateo",
                            "Santa Clara",
                            "Santa Cruz",
                            "Solano",
                            "Sonoma"))

ggplotly(bay_norm_total_log, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

Alameda
=====================================

### Confirmed Cases
```{r}
ac_tot_plot <- whisker.render(template_tot,ac_data)
eval(parse(text=ac_tot_plot))
```

### Additional cases by day

```{r}
ac_new_plot <- whisker.render(template_new,ac_data)
eval(parse(text=ac_new_plot))
```

### Deaths
```{r}
ac_deaths_plot <- whisker.render(template_deaths,ac_data)
eval(parse(text=ac_deaths_plot))
```

Contra Costa
=====================================

### Confirmed Cases
```{r}
cc_tot_plot <- whisker.render(template_tot,cc_data)
eval(parse(text=cc_tot_plot))
```

### Additional cases by day

```{r}
cc_new_plot <- whisker.render(template_new,cc_data)
eval(parse(text=cc_new_plot))
```

### Deaths
```{r}
cc_deaths_plot <- whisker.render(template_deaths,cc_data)
eval(parse(text=cc_deaths_plot))
```

Marin
=====================================

### Confirmed Cases
```{r}
mc_tot_plot <- whisker.render(template_tot,mc_data)
eval(parse(text=mc_tot_plot))
```

### Additional cases by day

```{r}
mc_new_plot <- whisker.render(template_new,mc_data)
eval(parse(text=mc_new_plot))
```

### Deaths
```{r}
mc_deaths_plot <- whisker.render(template_deaths,mc_data)
eval(parse(text=mc_deaths_plot))
```

Monterey
=====================================

### Confirmed Cases
```{r}
mo_tot_plot <- whisker.render(template_tot,mo_data)
eval(parse(text=mo_tot_plot))
```

### Additional cases by day

```{r}
mo_new_plot <- whisker.render(template_new,mo_data)
eval(parse(text=mo_new_plot))
```

### Deaths
```{r}
mo_deaths_plot <- whisker.render(template_deaths,mo_data)
eval(parse(text=mo_deaths_plot))
```


Napa
=====================================

### Confirmed Cases
```{r}
nc_tot_plot <- whisker.render(template_tot,nc_data)
eval(parse(text=nc_tot_plot))
```

### Additional cases by day

```{r}
nc_new_plot <- whisker.render(template_new,nc_data)
eval(parse(text=nc_new_plot))
```

### Deaths
```{r}
nc_deaths_plot <- whisker.render(template_deaths,nc_data)
eval(parse(text=nc_deaths_plot))
```

SF
=====================================

### Confirmed Cases
```{r}
sf_tot_plot <- whisker.render(template_tot,sf_data)
eval(parse(text=sf_tot_plot))
```

### Additional cases by day

```{r}
sf_new_plot <- whisker.render(template_new,sf_data)
eval(parse(text=sf_new_plot))
```

### Deaths
```{r}
sf_deaths_plot <- whisker.render(template_deaths,sf_data)
eval(parse(text=sf_deaths_plot))
```

San Mateo
=====================================

### Confirmed Cases
```{r}
sm_tot_plot <- whisker.render(template_tot,sm_data)
eval(parse(text=sm_tot_plot))
```

### Additional cases by day

```{r}
sm_new_plot <- whisker.render(template_new,sm_data)
eval(parse(text=sm_new_plot))
```

### Deaths
```{r}
sm_deaths_plot <- whisker.render(template_deaths,sm_data)
eval(parse(text=sm_deaths_plot))
```

Santa Clara
=====================================

### Confirmed Cases
```{r}
scc_tot_plot <- whisker.render(template_tot,scc_data)
eval(parse(text=scc_tot_plot))
```

### Additional cases by day

```{r}
scc_new_plot <- whisker.render(template_new,scc_data)
eval(parse(text=scc_new_plot))
```

### Deaths
```{r}
scc_deaths_plot <- whisker.render(template_deaths,scc_data)
eval(parse(text=scc_deaths_plot))
```

Santa Cruz
=====================================

### Confirmed Cases
```{r}
scz_tot_plot <- whisker.render(template_tot,scz_data)
eval(parse(text=scz_tot_plot))
```

### Additional cases by day

```{r}
scz_new_plot <- whisker.render(template_new,scz_data)
eval(parse(text=scz_new_plot))
```

### Deaths
```{r}
scz_deaths_plot <- whisker.render(template_deaths,scz_data)
eval(parse(text=scz_deaths_plot))
```

Solano
=====================================

### Confirmed Cases
```{r}
sol_tot_plot <- whisker.render(template_tot,sol_data)
eval(parse(text=sol_tot_plot))
```

### Additional cases by day

```{r}
sol_new_plot <- whisker.render(template_new,sol_data)
eval(parse(text=sol_new_plot))
```

### Deaths
```{r}
sol_deaths_plot <- whisker.render(template_deaths,sol_data)
eval(parse(text=sol_deaths_plot))
```


Sonoma
=====================================

### Confirmed Cases
```{r}
son_tot_plot <- whisker.render(template_tot,son_data)
eval(parse(text=son_tot_plot))
```

### Additional cases by day

```{r}
son_new_plot <- whisker.render(template_new,son_data)
eval(parse(text=son_new_plot))
```

### Deaths
```{r}
son_deaths_plot <- whisker.render(template_deaths,son_data)
eval(parse(text=son_deaths_plot))
```

More
=====================================

By [Jack J. Lee](https://www.jackjleescience.com)

Numbers from [L.A. Times coronavirus data](https://github.com/datadesk/california-coronavirus-data/blob/master/latimes-county-totals.csv).

Files available on [Github](https://github.com/jackjlee/covid19_bayarea_lat). Code last run on **`r Sys.Date()`**

Previous version of page archived at [https://www.jackjleescience.com/dataviz/covid19_bayarea_old.html](https://www.jackjleescience.com/dataviz/covid19_bayarea_old.html)